---
output:
  pdf_document: default
  html_document: default
---

```{r}
library(klaR)
library(caret)
library(rpart)
library(rsample)
library(dplyr)
library(rpart.plot)
library(tidyverse)
library(TTR)
library(randomForest)
library(corrplot)
library(httr)
library(jsonlite)

setwd('C:\\Users\\hiyam\\OneDrive\\Documents\\R Source\\P2')
set.seed(12345)
data <- read_csv('AAPL.csv')
glimpse(data)
nrow(data)

```

```{r}
#Initial plotting of AAPL close-prices
ggplot(data) + 
  geom_line(aes(x = Date, y = Close, color = "Original")) + 
  labs(title = "Time Series of AAPL Close Price", y = "Closings")
```

Feature Engineering : Calculating additional features from the initial Open, High, Close, Low features
```{r}

#calculate the RSI for a close price
calc_RSI <- function(price, n = 14) {
  return(RSI(price, n = n)/100); 
}

#This function turns trend strength into a vector quantity 
#When the dmi+ line is above the dmi-  line, prices are going up --> keep adx positive
#When the dmi- line is above the dmi+ line, prices are going down --> make adx negative
calc_trend_strength <- function(di_p, di_n, adx) {
    di_diff <- di_p - di_n
    strength <- ifelse((di_p - di_n) > 0, adx, -1 *adx)
    return(strength)
}

#When macd - signal is positive, macd > signal --> buy 
#when macd - signal is negative, macd < signal --> sell 
calc_macd_signal <- function(macd, signal) {
  return(macd - signal)
}

#bollinger bands calculation: SMA +- 2std
BB_matrix <- BBands(data$Close, n = 15, sd = 2, matype = "SMA")

#MACD calculation
macd_values <- MACD(data$Close, nFast = 12, nSlow = 26, nSig = 9, maType = "EMA", percent = FALSE)

#Stochastic Oscillator: make a matrix out of the high-low-close features
HLC <- data %>%
  select(High, Low, Close)

#Williams % Range
wpr <- WPR(HLC, n = 14)

#ADX/DMI+/DMI-
adx_values <- ADX(HLC, n = 14, maType = "EMA")
di_p <- adx_values[,"DIp"]
di_n <- adx_values[, "DIn"]
adx <- adx_values[, "ADX"]
```
Add columns for new features to original dataset
```{r}
#add columns for new features to original dataset
#created a column for price at day d - 1 (lagged_close) --> 1 = uptrend, 0 = downtrend

data <- data %>%
  mutate(rsi = calc_RSI(data$Close, n = 14),
         macd_signal = calc_macd_signal(macd_values[,"macd"], macd_values[,"signal"]), 
         macd = macd_values[, "macd"],
         signal = macd_values[,"signal"],
         will_pr = wpr,
         sma = SMA(data$Close, n = 14),
         lagged_close = lag(Close, 5),
         bb_up = BB_matrix[,"up"],
         bb_down = BB_matrix[,"dn"],
         trend_strength = calc_trend_strength(di_p, di_n, adx),
         direction = ifelse(Close > lagged_close, 1, 0))

data <- na.omit(data)

#Fix adj close col name
data <- data %>%
  mutate(adj_close = `Adj Close`) %>%
  select(-c(`Adj Close`))

print(data, n = 10, width = Inf) 

```
```{r}
close_data <- data.frame(data$Close, data$lagged_close, (data$Close - data$lagged_close))
close_data
```

Data Exploration and Plotting
```{r}
#Take a subset of the data
data_sub <- data[(nrow(data) - 100):(nrow(data)),]
print(data_sub, n = 20)
nrow(data_sub)

#MACD
ggplot(data_sub, aes(x = Date)) +
  geom_line(aes(y = macd, color = "MACD")) +
  geom_line(aes(y = signal, color = "Signal Line")) +
  labs(title = "MACD Indicator", y = "Value", color = "Legend")
#Close Prices
ggplot(data_sub, aes(x = Date)) + 
  geom_line(aes(y = Close, color = "Close")) + 
  labs(title = "Close Prices", y = "Close Price", color = "Legend")

#Examine how RSI works -- Close Price is scaled down so you can better see the patterns line up
#When RSI touches the 70 line, it indicates a stock being overbought
#When RSI touches the 30 line, it indicated a stock being oversold

#RSI
ggplot(data_sub, aes(x = Date)) + 
  geom_line(aes(y = rsi*100, color = "RSI")) +
  geom_line(aes(y = Close - 200, color = "Close")) + 
  geom_line(y = 70, color = "blue") + 
  geom_line(y = 30, color = "orange") +
  labs(title = "RSI", y = "Value", color = "Legend")

#ADX, DI
#Examine ADX, DI+, DI-
ggplot(data_sub, aes(x = Date)) + 
  geom_line(aes(y = trend_strength, color = "ADX")) +
  geom_line(aes(y = 0, color = "Base-Line")) + 
  geom_line(aes(y = -50, color = "Negative Strength")) + 
  geom_line(aes(y = 50, color = "Positive Strength")) + 
  labs(title = "ADX", y = "Value", color = "Legend")

#Williams %R
ggplot(data_sub, aes(x = Date)) + 
  geom_hline(yintercept = 20, color = "blue") + 
  geom_hline(yintercept = 80, color = "orange") +
  scale_y_reverse() + 
  geom_line(aes(y = will_pr * 100, color = "Will_PR")) + 
  labs(title = "Williams %R", y = "Value", color = "Legend")

```

```{r}
#Make all continuous features numeric and factorize target variable
data$rsi <- as.numeric(data$rsi)
data$macd <- as.numeric(data$macd)
data$signal <- as.numeric(data$signal)
data$will_pr <- as.numeric(data$will_pr)
data$sma <- as.numeric(data$sma)
data$bb_up <- as.numeric(data$bb_up)
data$bb_down <- as.numeric(data$bb_down)
data$trend_strength <- as.numeric(data$trend_strength)
data$direction <- as.factor(data$direction)

#Remove irrelevant features
data <- data %>%
  select(-c(Open, High, Low, Close, adj_close))

print(data, n = 15, width = Inf)
```
Examine correlation between features with a correlation matrix: correlations are expected
```{r}

cor_data <- data %>%
  select(-c(Date))
cor_matrix <- cor(cor_data[,-ncol(cor_data)])
print(cor_matrix)

corrplot(cor_matrix, method = "circle", type = "upper",
         tl.col = "black", tl.srt = 45,
         title = "Correlation of Stock Features", mar = c(0, 0, 1, 0))

```
Split data into training and test dataset without randomization because of time-series
```{r}
split_index <- floor(0.8 * nrow(data))
training <- data[1: split_index,]
test_data <- data[(split_index + 1):nrow(data), ]

prop.table(table(training$direction))
prop.table(table(test_data$direction))

head(training)
str(training)
```
RF Model with 100 trees, mtry is default
```{r}
rf_model <- randomForest(direction ~ rsi + macd + signal + will_pr + sma + bb_up + bb_down + trend_strength, data = training, ntree = 100, importance = TRUE)
print(rf_model)

pred <- predict(rf_model, test_data)
confusionMatrix(pred, test_data$direction)
importance(rf_model)
```






